---
title: "TPM calculation"
author: "Johannes Nicolaus Wibisana"
date: "11/4/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## TPM (transcripts per million) calculation

This tutorial covers transcripts per million calculation with library size normalization using edgeR.
This tutorial was constructed for The Laboratory of Cell Systems, Institute for Protein Research, Suita, Osaka, Japan.


prep our library
```{r message=FALSE}
library(tidyverse)
library(data.table)
library(edgeR)
```

let's read our raw counts data
```{r message=FALSE}
counts <- read_csv(file = "D:/lab_rmd/tpm/counts_trimmed.csv")

```

Here is how the data should look like

```{r paged.print=TRUE}
head(counts) 
```


Remove Chr, Start, End, Strand. Only keep geneid, length and samples

```{r}
counts <- counts[,!colnames(counts) %in% c("Chr", "Start", "End", "Strand")]
head(counts)
```


Divide the sample read count by the length of the gene multiplied by 1000 
```{r}
counts_tpm <- (counts[,3:ncol(counts)]/counts$Length) *1000
#return to tible just because I like tible
counts_tpm <- cbind(Geneid = counts$Geneid, counts_tpm) %>% as_tibble()

```


Get the normalization factor (in this case RLE)
```{r}
normfactor <- DGEList(counts = counts_tpm[,2:ncol(counts_tpm)],
                        group = colnames(counts_tpm[,2:ncol(counts_tpm)])
                        )

#you can change the normalization method here
normfactor <- calcNormFactors(normfactor, method="RLE")
  
normfactor_samples <- normfactor$samples
  
#multiply normalization factor with the library size
normfactor_samples$normlib <- normfactor_samples$lib.size*normfactor_samples$norm.factors
  
```


Loop to get the final TPM counts table
```{r}
for(i in 1:(dim(counts_tpm)[2]-1)){
    
  counts_tpm[,i+1] <- (counts_tpm[,i+1]/normfactor_samples$normlib[i])*1000000
    
}

```

```{r}
head(counts_tpm)

```



<!-- ## Comparison between normalized and unnormalized -->









<!-- You can also embed plots, for example: -->

<!-- ```{r pressure, echo=FALSE} -->
<!-- plot(pressure) -->
<!-- ``` -->

<!-- Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot. -->
